### 2020-09-21 parts/02-abstract.tex
Initial project template uploaded.

% A B S T R A C T
% ---------------

\begin{center}\textbf{Abstract}\end{center}

%This paper proposes a strategy for testing and comparing three control algorithms, LQR, LQG, and ADP, to control two two-DOF helicopters from a mobile device.  We will be using Raspberry Pi 3's as terminals for the wireless communication and MATLAB as our primary coding language.

This project proposes a modular and cost-effective smart real-time motion control
framework for a group of two degrees of freedom (2-DOF) helicopters.  The helicopters are
controlled by a mobile device which sends position signals over a wireless network to a
microcontroller.  The microcontroller uses control algorithms to determine the amount of voltage
to apply to DC motors.  This project tests and compares three control algorithms, LQR, LQG, and
ADP as well as theorizes possible methods to improve the control in future projects.

\cleardoublepage
%\newpage

%%% Local Variables:
%%% mode: latex
%%% TeX-master: "../finalReportMainV1"
%%% End: