### 2021-04-06 parts/03-acknowledgements.tex
Updated acknowledgements page

% A C K N O W L E D G E M E N T S
% -------------------------------

\begin{center}\textbf{Acknowledgements}\end{center}

    \par Special thanks to Mr. Christopher Mattus for his assistance in procuring the parts and modifying the hardware for our project.
    \par Thank you to everyone else who helped make this project possible and successful.

\cleardoublepage
%\newpage

%%% Local Variables:
%%% mode: latex
%%% TeX-master: "../finalReportMainV1"
%%% End: