### 2021-04-06 parts/02-abstract.tex

% A B S T R A C T
% ---------------

\begin{center}\textbf{Abstract}\end{center}

Robotic carts are of paramount importance for customers with disabilities in grocery stores. Although there are numerous robotic carts currently available, many times the hardware to implement the robot is quite costly. This project aims to develop a prototype of a smart robotic cart using radio frequency (RF) signals. The proposed robotic cart is cost-effective since it simply utilizes analog signal strengths to locate and follow the customer. An array of RF modules is mounted on top of the robotic cart, and a remote RF module is held by the customer. The direction and the line-of-sight (LoS) distance of the remote are estimated based on signal strengths received by the RF modules on the robot. The robot uses that direction and LoS distance to compute actuator commands for the robotic cart to follow the remote.  The robot was first simulated using CoppeliaSim, a commercial robot simulator. A prototype was then built and used to evaluate the performance of the robot.

\cleardoublepage
%\newpage

%%% Local Variables:
%%% mode: latex
%%% TeX-master: "../finalReportMainV1"
%%% End: