### 2021-04-07 parts/40-algorithms.tex

%======================================================================
\chapter{Algorithms}
\label{ch: Chapter4}
%======================================================================

%----------------------------------------------------------------------
\section{Robot Model}
%----------------------------------------------------------------------
The robot used in this project was a differential drive mobile robot.

%----------------------------------------------------------------------
\section{Localization Algorithm}
%----------------------------------------------------------------------
The localization algorithm is used to estimate the position of the remote with respect to the robot's local coordinate system.

%----------------------------------------------------------------------
\section{Navigation Algorithm}
%----------------------------------------------------------------------
The navigation algorithm is used to drive the robot toward the remote.

%%% Local Variables:
%%% mode: latex
%%% TeX-master: "../finalReportMainV1"
%%% End: