### 2021-04-13 parts/03-acknowledgements.tex
Update 03-acknowledgements.tex

% A C K N O W L E D G E M E N T S
% -------------------------------

\begin{center}\textbf{Acknowledgements}\end{center}

    \par Special thanks to Mr. Christopher Mattus for his assistance in procuring the parts and 
    \par modifying the hardware for our project.
    \par Special thanks to Shreel Patel for his assistance with charging the battery on the robot.
    \par Thank you to everyone else who helped make this project possible and successful.

\cleardoublepage
%\newpage

%%% Local Variables:
%%% mode: latex
%%% TeX-master: "../finalReportMainV1"
%%% End: