### 2021-05-01 parts/50-implementation.tex
Initial feedback from Dr. Miah

%======================================================================
\chapter{Implementation}
\label{ch: Chapter5}
%======================================================================

%----------------------------------------------------------------------
\section{Robot Assembly}
%----------------------------------------------------------------------

\todo[inline]{Please complete this section}

%----------------------------------------------------------------------
\section{Code Trial and Implementation}
%----------------------------------------------------------------------

\todo[inline]{Please complete this section}

%----------------------------------------------------------------------
\section{Experimental Results}
%----------------------------------------------------------------------

\todo[inline]{See page 23 of~\url{http://ee.bradley.edu/projects/proj2017/ekf_slam/Downloads/seniorProject_finalReport.pdf}}

\section{Simulation using Robot Simulator}

\todo[inline]{In this section, we detail the steps of simulating the proposed
  robotic cart using a commercial robot simulator CoppeliaSim. $\ldots \ldots$}

%%% Local Variables:
%%% mode: latex
%%% TeX-master: "../finalReport"
%%% End: