### 2021-05-01 parts/60-conclusion_and_future_work.tex

%======================================================================
\chapter{Conclusions and Future Work}
\label{ch: Chapter6}
%======================================================================

%----------------------------------------------------------------------
\section{Conclusions}
%----------------------------------------------------------------------

\todo[inline]{Complete the conclusion section}

%----------------------------------------------------------------------
\section{Future Work}
%----------------------------------------------------------------------

\todo[inline]{Complete the future work section}

%%% Local Variables:
%%% mode: latex
%%% TeX-master: "../finalReport"
%%% End: