### 2021-05-10 parts/40-conclusion_and_future_work.tex
Updated chapter labeling

%======================================================================
\chapter{Conclusion and Future Work}
\label{ch: conclusionAndFuture}
%======================================================================

%----------------------------------------------------------------------
\section{Conclusion}
%----------------------------------------------------------------------

\todo[inline]{Complete the conclusion section}

%----------------------------------------------------------------------
\section{Future Work}
%----------------------------------------------------------------------

\todo[inline]{Complete the future work section}

%%% Local Variables:
%%% mode: latex
%%% TeX-master: "../finalReport"
%%% End: